---
title: "Portfolio Piece 6: How Liberal Are the 2020 Democratic Candidates?"
author: "Rebecca Smith"
date: "4/26/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Leaning on ideological scores provided by OnTheIssues.org, this piece aims to provide evidence that the whole distribution of the 2020 Democratic candidates' ideologies is not that different from the positions of those who ran in 2008 or 2016. 

```{r load}
library(tidyverse)
library(ggalt)
library(ggthemes)
library(readr)
library(scales)
library(dplyr)
library(kableExtra)
cand_ideo <- read_csv("data/2019_02_08_how_liberal_2020_dems.csv")
```

First, we want to find the median scores along two dimensions of policies, 'economic' policies and 'social' policies for the past three presidential cycles with a Democratic primary. 

To do this, I'm first going to rescale the items and gather some summary stats.

```{r scale}
cand_ideo <- cand_ideo %>%
  mutate(economic = rescale(economic,c(-1,1),c(0,1)),
         social = rescale(social,c(-1,1),c(0,1)))
```

```{r summary}
cand_ideo %>%
  group_by(year) %>%
  summarise(social = median(social),
            economic = median(economic)) %>%
  kable()
options("stringsAsFactors" = TRUE)
```

And now, we plot. I did follow a guide provided by Elliott Morris for this visualization. 

```{r candideo}
cand_plot <- ggplot(cand_ideo, aes(x = economic, y = social, col=as.factor(year), fill=as.factor(year))) +
  geom_vline(xintercept = 0, linetype = 2) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(size=2) +
  coord_cartesian(xlim=c(-1,1),
                  ylim=c(-1,1)) +
  stat_bkde2d(bandwidth = c(0.12, 0.12), aes(alpha = (..level../1)*0.5), geom = "polygon", show.legend = F) +
  scale_color_wsj() +
  scale_fill_wsj() +
  guides("color"=guide_legend(title = "Year"),
         "fill" = "none") +
  labs(title = "2020 Democrats Look Ideologically Similar to their Predecessors",
       subtitle = "Coverage painting the 2020 candidates as record-setting liberals is unwarranted (Each point is a candidate.)",
       x = "Position on Economic Issues\n(-1=Liberal, 1=Conservative)",
       y = "Position on SOcial Issues\n(-1=Liberal, 1=Conservative)",
       caption="Source:OnTheIssues.org")

cand_plot
```

From this plot, we can see that the 2020 candidates' ideologies are significantly different than those who came before them. The common argument that Democrats are more liberal than ever before is not justified by the data. 